Surface programmable materials

### 2024-08-17 paper/main.tex
update

\section{Surface Programmable Materials}\label{surface-programmable-materials}

\subsection{Surface Programmable
Material}\label{surface-programmable-material}

@@

\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
Current Pattern (L, C, R) & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\
\hline
New State For Center Cell & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 \\
\hline
\end{tabular}

@@

We can encode the Rule 110 cellular automaton into a Weighted Maximum
Independent Set Problem, with blue vertices assigned a weight of 1 and
red vertices assigned a weight of 2, as follows.
\includegraphics[width=\columnwidth]{../notes/images/image.png}

This graph can be embedded into a grid graph, where two vertices are
connected if and only if their Euclidean distance is no more than $2$.
\includegraphics[width=\columnwidth]{../notes/images/image-1.png}

@@

The gadget we constructed based on the Rule 110 cellular automaton
naturally possesses Turing completeness. Therefore, it can be tiled in a
two-dimensional plane to create a computational material with logical
operation capabilities. Utilizing copy gadget and cross
gadget~\cite{Nguyen2023}, we construct a
\textbf{Surface Programmable Material} with open boundary conditions as
follows.

\begin{figure}
\centering
%\includesvg{../notes/images/rule110\_transvarient.svg}
\caption{Alt text}
\end{figure}

@@

\subsection{Deterministic and non-deterministic
computation}\label{deterministic-and-non-deterministic-computation}

Definitions: * \emph{in-surface/out-suface}: The surface of a surface
programmable material that associated with the input/output of the logic
circuit. \#\#\# Deterministic direction The computation contains the
following steps: 1. Initialize the in-surface configuration. By removing
some atoms on the in-surface. 2. Connect the in-surface/out-surface to
external heat sources at temperature $T_1 < T_2$, respectively. We
also require that the energy gap between the ground state and the first
excited state of the Hamiltonian to be $\Delta E< T_1$. 3. Lower the
temperature of the heat sources ``slowly'' to cool the system to the
ground state of the Hamiltonian. The temperature of the heat sources at
time $t$ is $T_{1/2}(t) = T_{1/2}(0)\lambda^{-c t}$, where
$T_{1/2}(0)$ is the initial temperature of the heat sources, and
$\alpha$ is a constant.

@@

\subsection{Outlook: The emergence of
wisdom}\label{outlook-the-emergence-of-wisdom}

@@

The gadget is dominated by such hamiltonian, correspond to the classical
part of Rydberg atoms.

\[
H = -\Delta \sum_i w_i \hat n_i + \sum_{|\vec r_i - \vec r_j|\leq 2} U \hat n_i \hat n_j
\]

Our target is to calculate the \textbf{state} with lowest energy under
the above hamiltonian as quickly as possible, while some positions are
set to be 1 or 0 in the \textbf{state}.

\[
H_{QAA}(t) = \sum_{v\in V} (-\Delta(t)w_v \hat n_v + \Omega(t)\sigma_{v}^x) + \sum_{(u,w) \in E} U\hat n_u \hat n_w
\]

@@

\[
H = \sum_{\text{each gadget}} [\text{gadget is not satisfied}]
\]

@@

where $h_i$ is onsite energy, $J_{u,v}$ is interaction energy and
$s_i \in {-1, 1}$ is local spin. The mapped spin-glass model should
possess propertys that it has exactly 8 ground states, each correspond
to one input-output relation in Cellular Automata.

@@

@@

>>>>

@@

\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
Current Pattern & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\
\hline
New Center Cell & 0 & 1 & 1 & 0 & 1 & 1 & 1 & 0 \\
\hline
\end{tabular}

@@

@@

Let us define a model Hamiltonian on a 2D lattice as
\begin{equation}
H = \sum_{\text{each gadget}} (n_{i+1, t} \land \neg(n_{i, t} \land \neg n_{i-1, t})) \veebar n_{i,t+1},
\end{equation}
where $n_{i, t} \in \{0, 1\}$ is the state of the cell at row $i$ and column $t$.
Its ground state with energy $0$ encodes a $110$ cellular automaton, where the direction of computation is from the left to right.

The gadget we constructed based on the Rule 110 cellular automaton
naturally possesses Turing completeness. Therefore, it can be tiled in a
two-dimensional plane to create a computational material with logical
operation capabilities.

\jgl{We need a main figure to show the tiling and cooling.}

@@

\subsection{The direction of computation}\label{deterministic-and-non-deterministic-computation}

Definitions: * \emph{in-surface/out-suface}: The surface of a surface
programmable material that associated with the input/output of the logic
circuit.

The computation contains the following steps: (1). Initialize the in-surface configuration. By removing
some atoms on the in-surface. (2). Connect the in-surface/out-surface to
external heat sources at temperature $T_1 < T_2$, respectively. We
also require that the energy gap between the ground state and the first
excited state of the Hamiltonian to be $\Delta E< T_1$. (3). Lower the
temperature of the heat sources ``slowly'' to cool the system to the
ground state of the Hamiltonian. The temperature of the heat sources at
time $t$ is $T_{1/2}(t) = T_{1/2}(0)\lambda^{-c t}$, where
$T_{1/2}(0)$ is the initial temperature of the heat sources, and
$\alpha$ is a constant.

@@

@@

The goal is to drive the the \textbf{state} to the one with lowest energy as fast as possible, while some bits set to be 1 or 0 in the \textbf{state}.

@@

@@

where $h_i$ is onsite energy, $J_{u,v}$ is interaction energy and
$s_i \in \{-1, 1\}$ is local spin. The mapped spin-glass model should
possess propertys that it has exactly 8 ground states, each correspond
to one input-output relation in Cellular Automata.

@@

\section{Discussion and Outlook}
Connect real-world thermalization time and the simulated annealing time to explore the connection between time and energy cost per computation.
The emergence of wisdom?

@@

We can encode the Rule 110 cellular automaton into a Weighted Maximum
Independent Set Problem, with blue vertices assigned a weight of 1 and
red vertices assigned a weight of 2, as follows.

\includegraphics[width=0.6\columnwidth]{../notes/images/image.png}

This graph can be embedded into a grid graph, where two vertices are
connected if and only if their Euclidean distance is no more than $2$.

\includegraphics[width=0.7\columnwidth]{../notes/images/image-1.png}

Utilizing copy gadget and cross gadget~\cite{Nguyen2023}, we construct a \textbf{Surface Programmable Material} with open boundary conditions as follows.

\begin{figure}
\centering
\includesvg[width=\columnwidth]{../notes/images/rule110transvarient.svg}
\caption{Alt text}
\end{figure}

\begin{equation}
H_{QAA}(t) = \sum_{v\in V} (-\Delta(t)w_v \hat n_v + \Omega(t)\sigma_{v}^x) + \sum_{(u,w) \in E} U\hat n_u \hat n_w
\end{equation}
